Estimation and Implications of Bias in EPA Pollution Measurement

### 2021-12-15 paper/5min-presentation/5min-presentation.Rmd
Made some chunks all appear at the same time.

---
title: "Estimation and Implications of Bias in EPA Pollution Measurement"
author: Aaron C Watt
date: December 2, 2021
output:
  beamer_presentation:
    includes:
      in_header: preamble.tex
classoption: "aspectratio=169"
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'acWatt_purpleair_5min-pres_2021-12-02.pdf')) })
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
# Had to run tinytex::install_tinytex() to use XeLaTeX if using YGbeamer.sty
```

@@

- The Clean Air Act (1970) established National Ambient Air Quality Standards
(NAAQS) for US counties

- Either "attainment" or "non-attainment", \color{red}large penalties and costly restrictions

- Minimum requirement of 75\% of readings, per quarter

- Air quality can change quickly

- Monitor shutoffs are common

@@

## Research Questions

- How biased are EPA monitor-based measures of local air quality?
<!-- - How biased are self-reported EPA pollution estimates? -->
<!-- - How biased is *missing* air pollution data from self-reporting US EPA monitors? -->

- Does this bias significantly change NAAQS attainment status?

@@

**Satellite Data:**

- Grainger et al. 2017
- Sullivan, Krupnick 2018
- Fowlie, Rubin, Walker 2019
- Zou 2021

**Location Pollution Alerts:**

- Mu, Rubin, Zou 2021

@@

## This project
- Using new consumer-based pollution monitors (PurpleAir).

- Avoids using satellite estimates (has been shown to have significant error).

- Focus on EPA pollution data that is missing *in time*, specifically PM2.5
<!-- - Understanding the distribution of bias in missing California air pollution data using consumer-based pollution -->
<!-- monitors that have low error near their location. -->

@@

load("/media/a/E/Programming/github/are219/data/epa/epa_monitors/epa_monitors/data_from_api/88101/api_request_daily_88101_2016.Rdata")
df = yr_dt %>% distinct(latitude, longitude, .keep_all = TRUE) %>% mutate(type=88101, NAAQS_used='yes')
load("/media/a/E/Programming/github/are219/data/epa/epa_monitors/epa_monitors/data_from_api/88501/api_request_daily_88501_2016.Rdata")
df2 = yr_dt %>% distinct(latitude, longitude, .keep_all = TRUE) %>% mutate(type=88501, NAAQS_used='no')
load("/media/a/E/Programming/github/are219/data/epa/epa_monitors/epa_monitors/data_from_api/88502/api_request_daily_88502_2016.Rdata")
df3 = yr_dt %>% distinct(latitude, longitude, .keep_all = TRUE) %>% mutate(type=88502, NAAQS_used='no')
epa_data = bind_rows(df, df2, df3) %>%
  filter(type == 88101) %>% 
  distinct(latitude, longitude, .keep_all = TRUE) %>%
  select(longitude, latitude, everything())

@@

```{r get PA sensor locations all, include=F}
result = jsonlite::fromJSON("/media/a/E/Programming/github/are219/data/test_data/manual_webpage_download_USmap_2021-11-29.txt")
library(zoo)
# Get start and last_seen dates from all sensors near the US
pa_data2 = as.data.frame(result$data) %>%
  purrr::set_names(result$fields) %>%
  mutate(date_start = lubridate::as_datetime(date_created),
         date_last = lubridate::as_datetime(last_seen),
         month_start = as.yearmon(date_start),
         month_last = as.yearmon(date_last)) %>%
  select(longitude, latitude, everything()) %>%
  arrange(date_start) %>%
  lonlat_country_filter(lon_col='longitude', lat_col='latitude') %>%
  filter(private == 0,  # public sensor
         location_type == 0,  # outdoor sensor
         channel_state != 0)  # has at least one PM sensor
```

@@

\begin{itemize}%[itemsep=1.5em]
\item \textbf{Estimate missing pollution observations:} Use PurpleAir data to predict EPA pollution at missing times
\item \textbf{Estimate bias of reported EPA pollution:} difference between predicted pollution at missing times and reported pollution at nonmissing times.
\item \textbf{Estimate counties' counterfactual attainment status:} Include estimated missing pollution data.
\end{itemize}

>>>>

---
title: "Estimation and Implications of Bias in EPA Pollution Measurement"
author: Aaron C Watt
date: December 2, 2021
output:
  beamer_presentation:
    includes:
      in_header: preamble.tex
classoption: "aspectratio=169"
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'acWatt_purpleair_5min-pres_2021-12-15.pdf')) })
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
# Had to run tinytex::install_tinytex() to use XeLaTeX if using YGbeamer.sty
```

@@

\begin{itemize}
\item<2-> The Clean Air Act (1970) established National Ambient Air Quality Standards
(NAAQS) for US counties
\item<2-> Either "attainment" or "non-attainment", \color{red}large penalties and costly restrictions
\item<2-> Minimum requirement of 75\% of readings, per quarter
\item<2-> Air quality can change quickly
\item<2-> Monitor shutoffs are common
\end{itemize}

@@

## Research Questions
\begin{itemize}
\item<3-> How biased are EPA monitor-based measures of local air quality?
\item<3-> Does this bias significantly change NAAQS attainment status?
\end{itemize}
<!-- - How biased are self-reported EPA pollution estimates? -->
<!-- - How biased is *missing* air pollution data from self-reporting US EPA monitors? -->

@@

\uncover<2->{**Satellite Data:**}
\begin{itemize}
\setlength{\itemsep}{0em}
\item<2-> Grainger et al. 2017
\item<2-> Sullivan, Krupnick 2018
\item<2-> Fowlie, Rubin, Walker 2019
\item<2-> Zou 2021
\end{itemize}

\uncover<3->{**Location Pollution Alerts:**
\begin{itemize}
\setlength{\itemsep}{0em}
\item<3-> Mu, Rubin, Zou 2021
\end{itemize}

@@

\uncover<4->{## This project}
\begin{itemize}
%\setlength{\itemsep}{0em}
\item<4-> Using new consumer-based pollution monitors (PurpleAir).
\item<4-> Avoids using satellite estimates (has been shown to have significant error).
\item<4-> Focus on EPA pollution data that is missing *in time*, specifically PM2.5
\end{itemize}

<!-- - Understanding the distribution of bias in missing California air pollution data using consumer-based pollution -->
<!-- monitors that have low error near their location. -->

@@

# The below data files are only on Aaron's desktop. Need to create another data submodule to host data
# or host on s3 bucket and download from there
load("../../data/epa/epa_monitors/epa_monitors/data_from_api/88101/api_request_daily_88101_2016.Rdata")
df = yr_dt %>% distinct(latitude, longitude, .keep_all = TRUE) %>% mutate(type=88101, NAAQS_used='yes')
load("../../data/epa/epa_monitors/epa_monitors/data_from_api/88501/api_request_daily_88501_2016.Rdata")
df2 = yr_dt %>% distinct(latitude, longitude, .keep_all = TRUE) %>% mutate(type=88501, NAAQS_used='no')
load("../../data/epa/epa_monitors/epa_monitors/data_from_api/88502/api_request_daily_88502_2016.Rdata")
df3 = yr_dt %>% distinct(latitude, longitude, .keep_all = TRUE) %>% mutate(type=88502, NAAQS_used='no')
epa_data = bind_rows(df, df2, df3) %>%
  filter(type == 88101) %>% 
  distinct(latitude, longitude, .keep_all = TRUE) %>%
  select(longitude, latitude, everything())

@@

```{r get PA sensor locations all, include=F}
result = jsonlite::fromJSON("../../data/test_data/manual_webpage_download_USmap_2021-11-29.txt")
library(zoo)
# Get start and last_seen dates from all sensors near the US
pa_data2 = as.data.frame(result$data) %>%
  purrr::set_names(result$fields) %>%
  mutate(date_start = lubridate::as_datetime(date_created),
         date_last = lubridate::as_datetime(last_seen),
         month_start = as.yearmon(date_start),
         month_last = as.yearmon(date_last)) %>%
  select(longitude, latitude, everything()) %>%
  arrange(date_start) %>%
  lonlat_country_filter(lon_col='longitude', lat_col='latitude') %>%
  filter(private == 0,  # public sensor
         location_type == 0,  # outdoor sensor
         channel_state != 0)  # has at least one PM sensor
```

@@

\begin{itemize}%[itemsep=1.5em]
\item<2> \textbf{Estimate missing pollution observations:} Use PurpleAir data to predict EPA pollution at missing times
\item<2> \textbf{Estimate bias of reported EPA pollution:} difference between predicted pollution at missing times and reported pollution at nonmissing times.
\item<2> \textbf{Estimate counties' counterfactual attainment status:} Include estimated missing pollution data.
\end{itemize}
